### 2016-04-18 true-grid/01_Abstract.tex
added a folder for SEMS

%!TEX root = 00_Paper_SEMS16.tex
\begin{abstract}
In the past years, we have seen several papers analyzing large spreadsheet corpora. Also, there is a growing body of research on the appliance of software engineering methods on spreadsheets. These studies focus to a large extent on formulas in spreadsheets. However, there are other interesting concepts in spreadsheets like Pivot Tables and charts. Users make use of these concepts to analyze and visualize large data sets. Despite the popularity of Pivot Tables and Charts, there is almost no research about their use. In this paper, we analyze two popular spreadsheet corpora: ENRON and EUSES on the use of Pivot Tables and charts.
\end{abstract}